---
title: "Analysis of Traffic Accidents in Zipcode 85257"
subtitle: "A Statistical Approach Using Regression and Estimation Techniques"
date: October 17, 2024
output: ioslides_presentation
---

<style type="text/css">
body p, div, h1, h2, h3, h4, h5 {
color: black !important;
font-family: Modern Computer Roman;
}
slides > slide.title-slide hgroup h1 {
color: black !important;
}
h2 {
color: black !important; 
}
</style>

```{r setup, include=FALSE}


library(ggplot2)
library(dplyr)
library(plotly)
library(tidyr)
data <- read.csv("US_Accidents_85257.csv")

total_accidents_85257 <- data %>% filter(Zipcode == "85257") %>% nrow()
filtered_data <- data %>% filter(Zipcode == "85257")
filtered_data$Start_Time <- as.POSIXct(filtered_data$Start_Time, format="%Y-%m-%d %H:%M:%S")
filtered_data_clean <- filtered_data %>% drop_na(Start_Time, Temperature.F., Severity)
accidents_per_day <- filtered_data_clean %>%
  mutate(Date = as.Date(Start_Time)) %>%
  group_by(Date) %>%
  summarize(Daily_Accidents = n())

lambda <- mean(accidents_per_day$Daily_Accidents)
```




## Introduction
- Over the time period from (February 2016 - March 2023) there has been a total of **`r total_accidents_85257`** Accidents in the 85257 Zipcode Recorded
- Using "A Countrywide Traffic Accident Dataset" from Sobhan Moosavi we will be analyzing patterns and relationships of accidents occurring in 85257
- Data set found here https://www.kaggle.com/datasets/sobhanmoosavi/us-accidents/data

## Exploration of The Data

- Average accident severity: **`r mean(filtered_data$Severity, na.rm = TRUE)`**
- Range of accident dates: **`r min(filtered_data$Start_Time)`** to **`r max(filtered_data$Start_Time)`**

```{r,echo=FALSE,fig.width=6, warning=FALSE}
plot_ly(filtered_data_clean, 
        x = ~Start_Time, 
        y = ~Temperature.F., 
        z = ~Severity, 
        type = "scatter3d", 
        mode = "markers", 
        marker = list(size = 3, color = ~Severity, colorscale = 'sunsetdark', showscale = TRUE),
  
        width = 600,
        height = 400
) %>%
  layout(
    title = "3D Scatter Plot of Accident Severity, Temperature, and Time of Accident in Zipcode 85257",
    scene = list(
      xaxis = list(title = 'Time of Accident',
                    range = c("2016-02-01", "2023-03-31")
      ),
      yaxis = list(title = 'Temperature (F)'),
      zaxis = list(title = 'Severity')
    )
  )

```


## The code to create the previous slide



<font size="1">
```{r, eval=FALSE}

plot_ly(filtered_data_clean, 
        x = ~Start_Time, 
        y = ~Temperature.F., 
        z = ~Severity, 
        type = "scatter3d", 
        mode = "markers", 
        marker = list(size = 3, color = ~Severity, colorscale = 'sunsetdark', showscale = TRUE),
  
        width = 600,
        height = 400
)

```
</font>

## The code to create the previous slide part 2: Layout
```{r,eval=FALSE}
layout(
    title = "3D Scatter Plot of Accident Severity, Temperature, and Time of Accident in Zipcode 85257",
    scene = list(
      xaxis = list(title = 'Time of Accident',
                    range = c("2016-02-01", "2023-03-31")
      ),
      yaxis = list(title = 'Temperature (F)'),
      zaxis = list(title = 'Severity')
    )
  )
```
## Linear Regression

- **Linear Regression between Temperature and Accident Severity**
```{r,echo=FALSE, message=FALSE, warning=FALSE}
ggplot(filtered_data_clean, aes(x = Temperature.F., y = Severity)) +
  geom_jitter(width = 0.3, height = 0.1) +
  geom_smooth(method = 'lm')
```


## Number of Accidents during Day vs Night

```{r, cars, echo=FALSE}
day_night_data <- filtered_data_clean %>%
  group_by(Sunrise_Sunset) %>%
  summarize(Accident_Count = n())


ggplot(day_night_data, aes(x = Sunrise_Sunset, y = Accident_Count, fill = Sunrise_Sunset)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Accident_Count), vjust = -0.5, size = 5) +
  labs(title = "Bar Plot of the Number of Accidents: Day vs. Night",
       x = "Time of Day",
       y = "Accident Count") +
  theme_dark() +
  scale_fill_manual(values = c("Day" = "blue", "Night" = "purple")) +
  theme(legend.position = "none")
```

## Important Severity Statisitcs

- **Mean Severity**: $\bar{x} = \frac{1}{n} \sum_{i=1}^{n} x_i$ = **`r mean(filtered_data_clean$Severity, na.rm = TRUE)`**
- **Variance**: $\sigma^2 = \frac{1}{n} \sum_{i=1}^{n} (x_i - \bar{x})^2$ = **`r var(filtered_data_clean$Severity, na.rm=TRUE)`**
- **Standard Deviation**: $\sigma = \sqrt{\frac{1}{n} \sum_{i=1}^{n} (x_i - \bar{x})^2}$ = **`r sd(filtered_data_clean$Severity, na.rm=TRUE)`**

## Poisson Distribution for Daily Accident Frequency

The probability of observing $k$ accidents in a fixed interval is given by the Poisson formula:

$$
P(X = k) = \frac{\lambda^k e^{-\lambda}}{k!}
$$

where:
- $\lambda$ is the average rate of accidents per day
- $k$ is the number of accidents observed


## Poisson Distribution for Daily Accident Frequency
- **Average accidents per day ($\lambda$): `r round(lambda, 2)`**

```{r,echo=FALSE,warning=FALSE}


poisson_data <- data.frame(
  x = 0:max(accidents_per_day$Daily_Accidents),
  Probability = dpois(0:max(accidents_per_day$Daily_Accidents), lambda)
)

ggplot(accidents_per_day, aes(x = Daily_Accidents)) +
  geom_histogram(aes(y = ..density..), binwidth = 1, fill = "blue", alpha = 0.5) +
  geom_line(data = poisson_data, aes(x = x, y = Probability), color = "red", linewidth = 1.3) +
  labs(
    title = "Graph of Actual vs. Poisson Distribution of Daily Accidents",
    x = "Number of Accidents per Day",
    y = "Density"
  ) +
  theme_dark()
```

